Multi-level generalizations of collective spin dynamics and squeezing

### 2019-10-03 collective_sun/sun_notes.tex
some cleanup in notations and conventions

We consider a collection of $n$-level fermions on a 1-D lattice, with
a single-particle Hamiltonian
\begin{align}
  H_{\t{lat}} = -J \sum_{q,\mu} \cos\p{qa} c_{q\mu}^\dag c_{q\mu},
  \label{eq:H_lat}
\end{align}
where $J$ is the nearest-neighbor tunneling rate, $a$ is the spacing
between neighboring lattice sites, and $c_{q\mu}$ is the fermionic
annihilation operator for a particle in quasi-momentum mode $q$ with
internal state (e.g.~nuclear spin) $\mu$.  Under the frozen-mode
approximation, two-body collective SU($n$)-symmetric interactions
between these fermions take form\cite{perlin2019effective}
\begin{align}
  H_{\t{int}} = \f{u}{2} \sum_{p,q,\mu,\nu}
  \p{c_{p\mu}^\dag c_{q\nu}^\dag c_{q\nu} c_{p\mu}
    + c_{q\mu}^\dag c_{p\nu}^\dag c_{q\nu} c_{p\mu}},
\end{align}
where $u\equiv U/L$ is the two-body on-site interaction energy $U$
divided by the number of lattice sites $L$.  Defining the spin
operators $s_{\mu\nu}^{(p)}\equiv c_{p\mu}^\dag c_{p\nu}$ for each
quasi-momentum $p$, we can alternately write
\begin{align}
  H_{\t{int}} = \f{u}{2} \sum_{p,q,\mu,\nu}
  \p{s_{\mu\mu}^{(p)} s_{\nu\nu}^{(q)} - s_{\mu\nu}^{(p)} s_{\nu\mu}^{(q)}}
  = \f{u}{2}\p{N^2 - \sum_{p,q,\mu,\nu} s_{\mu\nu}^{(p)} s_{\nu\mu}^{(q)}},
\end{align}
where $N$ is the total number of particles.  In the absence of
coherence between sectors of different particle number, we can neglect
the $\sim N^2$ term and simply write
\begin{align}
  H_{\t{int}}
  = - \f{u}{2} \sum_{p,q,\mu,\nu} s_{\mu\nu}^{(p)} s_{\nu\mu}^{(q)}.
\end{align}
In the case of SU(2), at this point we would expand the spin operators
$s_{\mu\nu}^{(p)}$ in terms of Pauli operators in order to convert
$H_{\t{int}}$ into an SU(2) spin Hamiltonian.  To generalize this
procedure to the case of SU($n$), we first define a vector
$\v s^{(p)}$ of all spin operators $s_{\mu\nu}^{(p)}$, and write
\begin{align}
  H_{\t{int}}
  = - \f{u}{2} \sum_{p,q} {\v s^{(p)}}^\dag \v s^{(q)}
  = - \f{u}{2} \v\S^\dag \c \v\S,
  &&
  \v\S\equiv \sum_p \v s^{(p)}.
  \label{eq:H_int_spin}
\end{align}
In the following sections, we identify various bases for the space of
operators on the Hilbert space of an $n$-level system.  For brevity,
we denote this space of operators by $\B_n$, and note that $\B_n$ is
itself an $n^2$-dimensional Hilbert space equipped with a trace
(Hilbert-Schmidt) inner product
\begin{align}
  \obk{\O|\Q} \equiv \tr\p{\O^\dag Q},
\end{align}
where $\oket{\Q}$ and $\obra{\O}$ respectively denote vectors in
$\B_n$ and its dual space $\B_n^*$.  Similarly to the set of all spin
operators $s_{\mu\nu}\equiv\op{\mu}{\nu}$ on the single-particle
Hilbert space at each quasi-momentum, all of the bases we consider
will be orthonormal with respect to the trace inner product.  As we
show in Appendix \ref{sec:changing_bases}, the collective
SU($n$)-symmetric interaction Hamiltonian in \eqref{eq:H_int_spin}
takes an identical form in any orthonormal basis for $\B_n$.

@@

An driving laser addressing spin-$1/2$ fermions on a 1-D lattice will
induce the Hamiltonian
\begin{align}
  \left. H_{\t{drive}}^{(\phi)} \right|_{n=2}
  = \Omega \sum_j
  \p{e^{-i\phi_j} c_{j,\up}^\dag c_{j,\dn}
    + e^{i\phi_j} c_{j,\dn}^\dag c_{j,\up}},
  \label{eq:H_drive_2}
\end{align}
where $\Omega$ is a (real) driving amplitude, $j$ indexes a lattice
site, and $\phi_j$ is a site-dependent phase.  The multi-level
analogue of \eqref{eq:H_drive_2} is
\begin{align}
  H_{\t{drive}}^{(\phi)}
  = -\Omega \sum_j \p{e^{-i\phi_j} T_{1,1}^{(j)}
    + e^{i\phi_j} {T_{1,1}^{(j)}}^\dag}
  = \Omega \D_{1,1}^{(\phi)},
  \label{eq:H_drive}
\end{align}
where
\begin{align}
  \D_{LM}^{(\phi)}
  \equiv \sum_j \sp{\cos\p{M\phi_j} D_{LM}^{(j)}
    + \sin\p{M\phi_j} D_{L,-M}^{(j)}},
  \label{eq:drive_rot}
\end{align}
is an inhomogeneously rotated collective drive operator.  We can make
this drive spatially homogenous via the gauge transformation
$c_{j\mu}^\dag \to e^{i\mu\phi_j} c_{j\mu}^\dag$, which takes
\begin{align}
  H_{\t{drive}}^{(\phi)} \to H_{\t{drive}} = \Omega \D_{1,1}.
\end{align}
The SU($n$) symmetry of collective interactions implies that the
interaction Hamiltonian $H_{\t{int}}$ in \eqref{eq:H_int_drive} is
invariant under this gauge transformation, but the single-particle
Hamiltonian in \eqref{eq:H_lat} transforms as
\begin{align}
  H_{\t{lat}}
  \to H_{\t{lat}}^{(\phi)}
  = -J \sum_{q,\mu} \cos\p{qa-\mu\phi} c_{q\mu}^\dag c_{q\mu},
  \label{eq:H_lat_SOC}
\end{align}
where $\phi$ is the relative phase of the driving laser between
adjacent lattice sites.  In the basis of drive operators, this
Hamiltonian takes the form (see Appendix \ref{sec:lat_drive})
\begin{align}
  H_{\t{lat}}^{(\phi)} = -J \sum_{q,L} B_L^{(\phi,q)} D_{L,0}^{(q)},
  \label{eq:H_lat_drive}
\end{align}
with effective inhomogeneous driving field strengths
\begin{align}
  B_L^{(\phi,q)} = f_L\p{qa} \sqrt{\f{L+1/2}{n}}
  \sum_\mu \bk{I\mu,L,0;I\mu} f_L\p{\mu\phi},
  &&
  f_L\p{\theta} \equiv
  \begin{cases}
    \cos\theta & L~\t{even} \\
    \sin\theta & L~\t{odd}
  \end{cases}.
\end{align}
Spin-orbit coupling with multi-level systems thus allows for the
simulation of synthetic gauge fields with large spin ($L>1$).  As we
will see in the following section, multi-laser drives can also
simulate large-spin synthetic fields with non-zero spin projection
($M\ne0$) onto a quantization axis.

@@

  = \sum_j \sum_{m\in\set{1,0,-1}} \p{\Omega_m e^{-i\phi_{jm}}

@@

amplitude; $\phi_{jm}$ is a site- and drive-dependent phase;
$\op{r}{s}$ for $r,s\in\set{\up,\dn}$ is a pseudo-spin operator for
the auxiliary degree of freedom off-resonantly addressed by the
lasers; $\Delta$ is a detuning of the lasers from the energy of an
excited state; and $\1$ is the identity operator.  If the detuning is
large, $\Delta\gg\Omega_m$, then at second order in perturbation
theory we can find that the effective action of
$H_{\t{Raman,bare}}^{(\phi)}$ within the $\ket{\dn}$-state manifold is

@@

  \f{\Omega_0\Omega_s}{\Delta} e^{-i\phi_{js}}

@@

  \sum_j\sum_{s\in\set{\pm1}} e^{-is\p{\phi_{j+}-\phi_{j-}}}

@@

where by choice of gauge we set $\phi_{j,0}=0$, or equivalently
re-define $\phi_{jm}-\phi_{j,0}\to\phi_{jm}$, and we identify the
subscripts $\pm\leftrightarrow\pm1$.  In terms of the drive operators,
we can expand

@@

phase-matched with $\phi_{js}=s\phi_j$, such that a particle on site
$j$ acquires a phase $\mp\phi_j$ for every nuclear spin transition

@@

  \left. H_{\t{Raman}}^{(\phi)} \right|_{\phi_{js}=s\phi_j}

@@

$c_{j\mu}^\dag \to e^{i\mu\phi_j} c_{j\mu}^\dag$, which takes

@@

Here we take the single-particle spin-orbit coupled Hamiltonian
$H_{\t{lat}}^{(\phi)}$ in \eqref{eq:H_lat_SOC}, and derive its
expression in \eqref{eq:H_lat_drive} in terms of the drive operators
$D_{LM}$.  In the basis of ordinary spin operators $s_{\mu\mu}$, this
Hamiltonian takes the form
\begin{align}
  H_{\t{lat}}^{(\phi)}
  = -J \sum_{q,\mu} \cos\p{qa-\mu\phi} s_{\mu\mu}^{(q)}
  \label{eq:H_lat_spin}
\end{align}
Defining the diagonal ``spin-wave'' operators
\begin{align}
  w_{\phi,+} \equiv \sum_\mu \cos\p{\mu\phi} s_{\mu\mu},
  &&
  w_{\phi,-} \equiv \sum_\mu \sin\p{\mu\phi} s_{\mu\mu},
\end{align}
which are respectively even and odd under spin inversion
($\mu\to-\mu$), we can simplify
\begin{align}
  \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
  = \cos\p{qa} w_{\phi,+} + \sin\p{qa} w_{\phi,-},
\end{align}
and resolve the identity (super-)operator $\I$ with
$\I w_{\phi,\pm} = w_{\phi,\pm}$ in the basis of the drive
operators $D_{LM}$, finding (see Appendix \ref{sec:changing_bases})
\begin{align}
  \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
  = \cos\p{qa} \f12 \sum_{L,M} \obk{w_{\phi,+}|D_{LM}} D_{LM}
  + \sin\p{qa} \f12 \sum_{L,M} \obk{w_{\phi,-}|D_{LM}} D_{LM}.
\end{align}
We now note that the drive operators $D_{LM}$ are strictly
off-diagonal for $M\ne0$, and furthermore that $D_{L,0}$ with even
(odd) $L$ is even (odd) under spin inversion, which implies
\begin{align}
  \sum_{L,M} \obk{w_{\phi,+}|D_{LM}} D_{LM}
  = \sum_{L\,\t{even}} \obk{w_{\phi,+}|D_{L,0}} D_{L,0},
\end{align}
and likewise with $w_{\phi,-}$ and odd $L$.  In total, we can
write
\begin{align}
  \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
  = \cos\p{qa} \f12 \sum_{L\,\t{even}}
  \obk{w_{\phi,+}|D_{L,0}} D_{L,0}
  + \sin\p{qa} \f12 \sum_{L\,\t{odd}}
  \obk{w_{\phi,-}|D_{L,0}} D_{L,0},
\end{align}
or, more compactly,
\begin{align}
  \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
  = \sum_L B_L^{(\phi,q)} D_{L,0},
  \label{eq:compact_SOC_drive}
\end{align}
with effective inhomogeneous driving field strengths
\begin{align}
  B_L^{(\phi,q)}
  \equiv f_L\p{qa} \f12
  \sum_\mu f_L\p{\mu\phi} \obk{s_{\mu\mu}|D_{L,0}}
  = f_L\p{qa} \sqrt{\f{L+1/2}{n}}
  \sum_\mu \bk{I\mu,L,0;I\mu} f_L\p{\mu\phi},
\end{align}
where
\begin{align}
  f_L\p{\theta} \equiv
  \begin{cases}
    \cos\theta & L~\t{even} \\
    \sin\theta & L~\t{odd}
  \end{cases}.
\end{align}
Substituting the result in \eqref{eq:compact_SOC_drive} into the
single-particle Hamiltonian in \eqref{eq:H_lat_spin} yields
\begin{align}
  H_{\t{lat}}^{(\phi)} = -J \sum_{q,L} B_L^{(\phi,q)} D_{L,0}^{(q)}.
\end{align}

@@

Here we provide the coefficients $\Omega_{LM}^{(\phi,j)}$ of the
effective Raman driving Hamiltonian
\begin{align}
  H_{\t{Raman}}^{(\phi)}
  = \sum_j \sum_{\substack{L\in\set{0,1,2}\\\abs{M}\le L}}
  \Omega_{LM}^{(\phi,j)} D_{LM}^{(j)}
\end{align}
considered in Section \ref{sec:drive_raman}.  These coefficients are
computed by evaluating products of transition operators in
\eqref{eq:raman_0}--\eqref{eq:raman_2}, using both the structure
constants of the transition operator algebra (derived in Appendix
\ref{sec:trans_prod}) and the fact that
$T_{LM}^\dag = \p{-1}^M T_{L,-M}$.  For clarity, if any coefficient
$\Omega_{LM}^{(\phi,j)}$ is independent of $\phi$, we drop the
explicit dependence on $\phi$, i.e.~taking
$\Omega_{LM}^{(\phi,j)}\to\Omega_{LM}$.
\begin{align}
  \Omega_{0,0} \equiv -\f1{\sqrt{2n}} \sum_m\f{\Omega_m^2}{\Delta},
  &&
  \Omega_{1,0} \equiv \sqrt{\f{3}{2n\p{n^2-1}}}~
  \f{\Omega_+^2 - \Omega_-}{\Delta},
  \label{eq:O_01_0}
\end{align}
\begin{align}
  \Omega_{2,0} \equiv \sqrt{\f{n^2-4}{10n\p{n^2-1}}}~
  \f{\Omega_+ + \Omega_- - 2\Omega_0}{\Delta},
  \label{eq:O_2_0}
\end{align}
\begin{align}
  \Omega_{L,1}^{(\phi,j)}
  \equiv \sum_{s\in\set{\pm1}} \omega_{Ls} \cos\phi_{js},
  &&
  \Omega_{L,-1}^{(\phi,j)}
  \equiv \sum_{s\in\set{\pm1}} s \omega_{Ls} \sin\phi_{js},
\end{align}
\begin{align}
  \Omega_{2,2}^{(\phi,j)}
  \equiv \omega_{2,2} \cos\p{\phi_{j+}-\phi_{j-}},
  &&
  \Omega_{2,-2}^{(\phi,j)}
  \equiv \omega_{2,2} \sin\p{\phi_{j+}-\phi_{j-}},
\end{align}
\begin{align}
  \omega_{1\pm} \equiv \f{\sqrt3}{\sqrt{n\p{n^2-1}}}~
  \f{\Omega_0 \Omega_\pm}{\Delta},
  &&
  \omega_{2\pm} \equiv \pm \f{\sqrt{3/5}}{n\p{n^2-1}}
  \sqrt{\f{\p{n+2}!}{\p{n-3}!}}~ \f{\Omega_0\Omega_\pm}{\Delta},
\end{align}
\begin{align}
  \omega_{2,2} \equiv -\f{\sqrt{6/5}}{n\p{n^2-1}}
  \sqrt{\f{\p{n+2}!}{\p{n-3}!}}~ \f{\Omega_+\Omega_-}{\Delta}.
\end{align}
Here $\Omega_m$, $\phi_{jm}$, and $\Delta$ are respectively the (real)
driving amplitudes, phases, and detuning that appear in the ``bare''
Raman drive in \eqref{eq:raman_bare}, and by choice of gauge we set
$\phi_{j,0}=0$, or equivalently re-define
$\phi_{jm}-\phi_{j,0}\to\phi_{jm}$.  If the driving lasers are
phase-matched with $\phi_{js}=s\phi_j$, such that a particle on site
$j$ acquires a phase $\mp\phi_j$ for every nuclear spin transition
$\mu\to\mu\pm1$, then the drive $H_{\t{Raman}}^{(\phi)}$ takes the
form
\begin{align}
  \left. H_{\t{Raman}}^{(\phi)} \right|_{\phi_{js}=s\phi_j}
  = \sum_{\substack{L\le2\\0\le M\le L}} \Omega_{LM} \D_{LM}^{(\phi)},
\end{align}
with inhomogeneously rotated collective drive operators
\begin{align}
  \D_{LM}^{(\phi)}
  \equiv \sum_j \sp{\cos\p{M\phi_j} D_{LM}^{(j)}
    + \sin\p{M\phi_j} D_{L,-M}^{(j)}},
\end{align}
and uniform coefficients given by \eqref{eq:O_01_0}, \eqref{eq:O_2_0},
and
\begin{align}
  \Omega_{1,-1} = \Omega_{2,-1} = \Omega_{2,-2} \equiv 0,
  &&
  \Omega_{L,1} \equiv \sum_{s\in\set{\pm1}} \omega_{Ls},
  &&
  \Omega_{2,2} \equiv \omega_{2,2}.
\end{align}

>>>>

We consider a collection of $n$-level fermions on a 1-D lattice, with
a single-particle Hamiltonian
\begin{align}
  H_{\t{lat}} = -J \sum_{q,\mu} \cos\p{qa} c_{q\mu}^\dag c_{q\mu},
  \label{eq:H_lat}
\end{align}
where $J$ is the nearest-neighbor tunneling rate, $a$ is the spacing
between neighboring lattice sites, and $c_{q\mu}$ is the fermionic
annihilation operator for a particle in quasi-momentum mode $q$ with
internal state (e.g.~nuclear spin) $\mu$.  Under the frozen-mode
approximation, two-body collective SU($n$)-symmetric interactions
between these fermions take form\cite{perlin2019effective}
\begin{align}
  H_{\t{int}} = \f{u}{2} \sum_{p,q,\mu,\nu}
  \p{c_{p\mu}^\dag c_{q\nu}^\dag c_{q\nu} c_{p\mu}
    + c_{q\mu}^\dag c_{p\nu}^\dag c_{q\nu} c_{p\mu}},
\end{align}
where $u\equiv U/L$ is the two-body on-site interaction energy $U$
divided by the number of lattice sites $L$.  Defining the spin
operators $s_{\mu\nu}^{(p)}\equiv c_{p\mu}^\dag c_{p\nu}$ for each
quasi-momentum $p$, we can alternately write
\begin{align}
  H_{\t{int}} = \f{u}{2} \sum_{p,q,\mu,\nu}
  \p{s_{\mu\mu}^{(p)} s_{\nu\nu}^{(q)} - s_{\mu\nu}^{(p)} s_{\nu\mu}^{(q)}}
  = \f{u}{2}\p{N^2 - \sum_{p,q,\mu,\nu} s_{\mu\nu}^{(p)} s_{\nu\mu}^{(q)}},
\end{align}
where $N$ is the total number of particles.  In the absence of
coherence between sectors of different particle number, we can neglect
the $\sim N^2$ term and simply write
\begin{align}
  H_{\t{int}}
  = - \f{u}{2} \sum_{p,q,\mu,\nu} s_{\mu\nu}^{(p)} s_{\nu\mu}^{(q)}.
\end{align}
In the case of SU(2), at this point we would expand the spin operators
$s_{\mu\nu}^{(p)}$ in terms of Pauli operators in order to convert
$H_{\t{int}}$ into an SU(2) spin Hamiltonian.  To generalize this
procedure to the case of SU($n$), we first define a vector
$\v s^{(p)}$ of all spin operators $s_{\mu\nu}^{(p)}$, and write
\begin{align}
  H_{\t{int}}
  = - \f{u}{2} \sum_{p,q} {\v s^{(p)}}^\dag \v s^{(q)}
  = - \f{u}{2} \v\S^\dag \c \v\S,
  &&
  \v\S\equiv \sum_p \v s^{(p)}.
  \label{eq:H_int_spin}
\end{align}
In the remainder of this section, we identify various bases for the
space of operators on the Hilbert space of an $n$-level system.  For
brevity, we denote this space of operators by $\B_n$, and note that
$\B_n$ is itself an $n^2$-dimensional Hilbert space equipped with a
trace (Hilbert-Schmidt) inner product
\begin{align}
  \obk{\O|\Q} \equiv \tr\p{\O^\dag Q},
\end{align}
where $\oket{\Q}$ and $\obra{\O}$ respectively denote vectors in
$\B_n$ and its dual space $\B_n^*$.  Similarly to the set of all spin
operators $s_{\mu\nu}\equiv\op{\mu}{\nu}$ on the single-particle
Hilbert space at each quasi-momentum, all of the bases we consider
will be orthonormal with respect to the trace inner product.  As we
show in Appendix \ref{sec:changing_bases}, the collective
SU($n$)-symmetric interaction Hamiltonian in \eqref{eq:H_int_spin}
takes an identical form in any orthonormal basis for $\B_n$.

@@

An driving laser addressing spin-$1/2$ fermions on a 1-D lattice will
induce the Hamiltonian
\begin{align}
  \left. H_{\t{drive}}^{(\phi)} \right|_{n=2}
  = \Omega \sum_j
  \p{e^{-i\phi j} c_{j,\up}^\dag c_{j,\dn}
    + e^{i\phi j} c_{j,\dn}^\dag c_{j,\up}},
  \label{eq:H_drive_2}
\end{align}
where $\Omega$ is a (real) driving amplitude, $j\in\mathbb{Z}$ indexes
a lattice site, and $\phi$ is the relative phase of the drive between
adjacent lattice sites.  The multi-level analogue of
\eqref{eq:H_drive_2} is
\begin{align}
  H_{\t{drive}}^{(\phi)}
  = -\Omega \sum_j \p{e^{-i\phi j} T_{1,1}^{(j)}
    + e^{i\phi j} {T_{1,1}^{(j)}}^\dag}
  = \Omega \D_{1,1}^{(\phi)},
  \label{eq:H_drive}
\end{align}
where
\begin{align}
  \D_{LM}^{(\phi)}
  \equiv \sum_j \sp{\cos\p{M\phi j} D_{LM}^{(j)}
    + \sin\p{M\phi j} D_{L,-M}^{(j)}},
  \label{eq:drive_rot}
\end{align}
is an inhomogeneously rotated collective drive operator.  We can make
this drive spatially homogenous via the gauge transformation
$c_{j\mu}^\dag \to e^{i\mu\phi j} c_{j\mu}^\dag$, which takes
\begin{align}
  H_{\t{drive}}^{(\phi)} \to H_{\t{drive}} = \Omega \D_{1,1}.
\end{align}
The SU($n$) symmetry of collective interactions implies that the
interaction Hamiltonian $H_{\t{int}}$ in \eqref{eq:H_int_drive} is
invariant under this gauge transformation, but the single-particle
Hamiltonian in \eqref{eq:H_lat} transforms as (see Appendix
\ref{sec:lat_drive})
\begin{align}
  H_{\t{lat}}
  \to H_{\t{lat}}^{(\phi)}
  = -J \sum_{q,\mu} \cos\p{qa-\mu\phi} c_{q\mu}^\dag c_{q\mu}
  = \sum_{q,L} B_L^{(\phi,q)} D_{L,0}^{(q)},
  \label{eq:H_lat_SOC}
\end{align}
with effective inhomogeneous driving field amplitudes
\begin{align}
  B_L^{(\phi,q)} = -J f_L\p{qa} \sqrt{\f{L+1/2}{n}}
  \sum_\mu \bk{I\mu,L,0;I\mu} f_L\p{\mu\phi},
  &&
  f_L\p{\theta} \equiv
  \begin{cases}
    \cos\theta & L~\t{even} \\
    \sin\theta & L~\t{odd}
  \end{cases}.
\end{align}
Spin-orbit coupling with multi-level systems thus allows for the
simulation of synthetic gauge fields with large spin ($L>1$).  As we
will see in the following section, multi-laser drives can also
simulate synthetic large-spin fields with non-zero spin projection
($M\ne0$) onto a quantization axis.

@@

  = \sum_j \sum_{m\in\set{1,0,-1}} \p{\Omega_m e^{-i\phi_m j}

@@

amplitude; $\phi_m$ is the relative phase of drive $m$ between
adjacent lattice sites; $\op{r}{s}$ for $r,s\in\set{\up,\dn}$ is a
pseudo-spin operator for the auxiliary degree of freedom
off-resonantly addressed by the lasers; $\Delta$ is a detuning of the
lasers from the energy of an excited state; and $\1$ is the identity
operator.  If the detuning is large, $\Delta\gg\Omega_m$, then at
second order in perturbation theory we can find that the effective
action of $H_{\t{Raman,bare}}^{(\phi)}$ within the $\ket{\dn}$-state
manifold is

@@

  \f{\Omega_0\Omega_s}{\Delta} e^{-i\phi_s j}

@@

  \sum_j\sum_{s\in\set{\pm1}} e^{-is\p{\phi_+-\phi_-}j}

@@

where by choice of gauge we set $\phi_0=0$, or equivalently re-define
$\phi_m-\phi_0\to\phi_m$, and we identify the subscripts
$\set{\pm}\leftrightarrow\set{\pm1}$.  In terms of the drive
operators, we can expand

@@

phase-matched with $\phi_s=s\phi$, such that a particle on site $j$
acquires a phase $\mp\phi j$ for every nuclear spin transition

@@

  \left. H_{\t{Raman}}^{(\phi)} \right|_{\phi_s=s\phi}

@@

$c_{j\mu}^\dag \to e^{i\mu\phi j} c_{j\mu}^\dag$, which takes

@@

Here we derive the drive-operator ($D_{LM}$) expansion of the
spin-orbit coupled Hamiltonian $H_{\t{lat}}^{(\phi)}$ in
\eqref{eq:H_lat_SOC}.  In the basis of spin operators
$s_{\mu\mu}^{(q)} \equiv c_{q\mu}^\dag c_{q\mu}$, this Hamiltonian
takes the form
\begin{align}
  H_{\t{lat}}^{(\phi)}
  = -J \sum_{q,\mu} \cos\p{qa-\mu\phi} s_{\mu\mu}^{(q)}
  \label{eq:H_lat_spin}
\end{align}
Defining the diagonal ``spin-wave'' operators
\begin{align}
  w_{\phi+} \equiv \sum_\mu \cos\p{\mu\phi} s_{\mu\mu},
  &&
  w_{\phi-} \equiv \sum_\mu \sin\p{\mu\phi} s_{\mu\mu},
\end{align}
which are respectively even and odd under spin inversion
($\mu\to-\mu$), we can expand
\begin{align}
  \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
  = \cos\p{qa} w_{\phi+} + \sin\p{qa} w_{\phi-},
\end{align}
and resolve the identity (super-)operator $\I$ with
$\I w_{\phi,\pm} = w_{\phi,\pm}$ in the basis of the drive
operators $D_{LM}$, finding (see Appendix \ref{sec:changing_bases})
\begin{align}
  \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
  = \cos\p{qa} \f12 \sum_{L,M} \obk{w_{\phi+}|D_{LM}} D_{LM}
  + \sin\p{qa} \f12 \sum_{L,M} \obk{w_{\phi-}|D_{LM}} D_{LM}.
\end{align}
We now note that the drive operators $D_{LM}$ are strictly
off-diagonal for $M\ne0$, and furthermore that $D_{L,0}$ with even
(odd) $L$ is even (odd) under spin inversion, which implies
\begin{align}
  \sum_{L,M} \obk{w_{\phi+}|D_{LM}} D_{LM}
  = \sum_{L\,\t{even}} \obk{w_{\phi+}|D_{L,0}} D_{L,0},
\end{align}
and likewise with $w_{\phi-}$ and odd $L$.  In total, we can
write
\begin{align}
  \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
  = \cos\p{qa} \f12 \sum_{L\,\t{even}}
  \obk{w_{\phi+}|D_{L,0}} D_{L,0}
  + \sin\p{qa} \f12 \sum_{L\,\t{odd}}
  \obk{w_{\phi,-}|D_{L,0}} D_{L,0},
\end{align}
or, more compactly,
\begin{align}
  \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
  = \sum_L A_L^{(\phi,q)} D_{L,0},
  \label{eq:compact_SOC_drive}
\end{align}
with effective inhomogeneous driving field amplitudes
\begin{align}
  A_L^{(\phi,q)}
  \equiv f_L\p{qa} \f12
  \sum_\mu f_L\p{\mu\phi} \obk{s_{\mu\mu}|D_{L,0}}
  = f_L\p{qa} \sqrt{\f{L+1/2}{n}}
  \sum_\mu \bk{I\mu,L,0;I\mu} f_L\p{\mu\phi},
\end{align}
where
\begin{align}
  f_L\p{\theta} \equiv
  \begin{cases}
    \cos\theta & L~\t{even} \\
    \sin\theta & L~\t{odd}
  \end{cases}.
\end{align}
Substituting the result in \eqref{eq:compact_SOC_drive} into the
single-particle Hamiltonian in \eqref{eq:H_lat_spin} yields
\begin{align}
  H_{\t{lat}}^{(\phi)} = \sum_{q,L} B_L^{(\phi,q)} D_{L,0}^{(q)},
  &&
  B_L^{(\phi,q)} \equiv -J A_L^{(\phi,q)}.
\end{align}

@@

Here we provide the coefficients $\Omega_{LM}^{(\phi,j)}$ of the
effective Raman driving Hamiltonian
\begin{align}
  H_{\t{Raman}}^{(\phi)}
  = \sum_j \sum_{\substack{L\in\set{0,1,2}\\\abs{M}\le L}}
  \Omega_{LM}^{(\phi,j)} D_{LM}^{(j)}
\end{align}
considered in Section \ref{sec:drive_raman}.  These coefficients are
computed by evaluating products of transition operators in
\eqref{eq:raman_0}--\eqref{eq:raman_2}, using both the structure
constants of the transition operator algebra (derived in Appendix
\ref{sec:trans_prod}) and the fact that
$T_{LM}^\dag = \p{-1}^M T_{L,-M}$.  For clarity, if any coefficient
$\Omega_{LM}^{(\phi,j)}$ is independent of $\phi$, we drop the
explicit dependence on $\phi$, i.e.~taking
$\Omega_{LM}^{(\phi,j)}\to\Omega_{LM}$.
\begin{align}
  \Omega_{0,0} \equiv -\f1{\sqrt{2n}} \sum_m\f{\Omega_m^2}{\Delta},
  &&
  \Omega_{1,0} \equiv \sqrt{\f{3}{2n\p{n^2-1}}}~
  \f{\Omega_+^2 - \Omega_-}{\Delta},
  \label{eq:O_01_0}
\end{align}
\begin{align}
  \Omega_{2,0} \equiv \sqrt{\f{n^2-4}{10n\p{n^2-1}}}~
  \f{\Omega_+ + \Omega_- - 2\Omega_0}{\Delta},
  \label{eq:O_2_0}
\end{align}
\begin{align}
  \Omega_{L,1}^{(\phi,j)}
  \equiv \sum_{s\in\set{\pm1}} \omega_{Ls} \cos\p{\phi_s j},
  &&
  \Omega_{L,-1}^{(\phi,j)}
  \equiv \sum_{s\in\set{\pm1}} s \omega_{Ls} \sin\p{\phi_s j},
\end{align}
\begin{align}
  \Omega_{2,2}^{(\phi,j)}
  \equiv \omega_{2,2} \cos\p{\sp{\phi_+-\phi_-}j},
  &&
  \Omega_{2,-2}^{(\phi,j)}
  \equiv \omega_{2,2} \sin\p{\sp{\phi_+-\phi_-}j},
\end{align}
\begin{align}
  \omega_{1\pm} \equiv \f{\sqrt3}{\sqrt{n\p{n^2-1}}}~
  \f{\Omega_0 \Omega_\pm}{\Delta},
  &&
  \omega_{2\pm} \equiv \pm \f{\sqrt{3/5}}{n\p{n^2-1}}
  \sqrt{\f{\p{n+2}!}{\p{n-3}!}}~ \f{\Omega_0\Omega_\pm}{\Delta},
\end{align}
\begin{align}
  \omega_{2,2} \equiv -\f{\sqrt{6/5}}{n\p{n^2-1}}
  \sqrt{\f{\p{n+2}!}{\p{n-3}!}}~ \f{\Omega_+\Omega_-}{\Delta}.
\end{align}
Here $\Omega_m$, $\phi_m$, and $\Delta$ are respectively the (real)
driving amplitudes, phases, and detuning that appear in the ``bare''
Raman drive in \eqref{eq:raman_bare}, and by choice of gauge we set
$\phi_-=0$, or equivalently re-define $\phi_m-\phi_0\to\phi_m$.  If
the driving lasers are phase-matched with $\phi_s=s\phi$, such that a
particle on site $j$ acquires a phase $\mp\phi j$ for every nuclear
spin transition $\mu\to\mu\pm1$, then the drive
$H_{\t{Raman}}^{(\phi)}$ takes the form
\begin{align}
  \left. H_{\t{Raman}}^{(\phi)} \right|_{\phi_s=s\phi}
  = \sum_{\substack{L\le2\\0\le M\le L}} \Omega_{LM} \D_{LM}^{(\phi)},
\end{align}
with inhomogeneously rotated collective drive operators
\begin{align}
  \D_{LM}^{(\phi)}
  \equiv \sum_j \sp{\cos\p{M\phi j} D_{LM}^{(j)}
    + \sin\p{M\phi j} D_{L,-M}^{(j)}},
\end{align}
and uniform coefficients given by \eqref{eq:O_01_0}, \eqref{eq:O_2_0},
and
\begin{align}
  \Omega_{1,-1} = \Omega_{2,-1} = \Omega_{2,-2} \equiv 0,
  &&
  \Omega_{L,1} \equiv \sum_{s\in\set{\pm1}} \omega_{Ls},
  &&
  \Omega_{2,2} \equiv \omega_{2,2}.
\end{align}
